### 2015-04-14 results.tex

\section*{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig1}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:pca}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig2}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:decline}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig3}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:heterotic}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig4}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:genphys}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig5}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:genphys2}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE