### 2015-07-31 remap_table1/remap.rmd
Updated genetic positions of Tbl_1 in main.tex

---
output: 
  html_document: 
    highlight: tango
    theme: cosmo
---
#Mapping the fixed haplotype regions to genetic coordinates.

```{r loadup, message=FALSE, warning=FALSE}
rm(list=ls())
library(dplyr)
snps = load("../data/new_full_set.RData")

#turn the dataset object into a dataframe
df <- data.frame(Chrom = new_full_set$LG,
                 Position = new_full_set$position,
                 Name = as.character(new_full_set$markernames),
                 Code = as.character(new_full_set$markercodes),
                 Gen = new_full_set$F2map,
                 stringsAsFactors = F)
```
##BSSS3.1
```{r}
chr3 = filter(df, Chrom==3)
head(chr3[order(abs(67.7e6-chr3$Position)),])
```
###Starting cM: 53.6 
```{r}
head(chr3[order(abs(chr3$Position-123e6)),], n=10)
```
###Ending cM:  55.3

##BSSS3.2
```{r}
head(chr3[order(abs(129.2e6-chr3$Position)),])
```
###Starting cM: 57.3
```{r}
head(chr3[order(abs(157.1e6-chr3$Position)),], n=10)
```
###Ending cM: 64.8

##BSSS4
```{r}
chr4 = filter(df, Chrom==4)
head(chr4[order(abs(39.9e6-chr4$Position)),],n=10)
```
###Starting cM: 52.8
```{r}
head(chr4[order(abs(82.7e6-chr4$Position)),],n=10)
```
###Ending cM: 55.6

##BSSS9.1
```{r}
chr9 = filter(df, Chrom==9)
head(chr9[order(abs(20.8e6-chr9$Position)),],n=10)
```
###Starting cM: 41.1
```{r}
head(chr9[order(abs(26.6e6-chr9$Position)),],n=10)
```
###Ending cM: 44.8

##BSSS9.2
```{r}
head(chr9[order(abs(30.8e6-chr9$Position)),],n=10)
```
###Starting cM: 45.7
```{r}
head(chr9[order(abs(90.4e6-chr9$Position)),],n=10)
```
###Ending cM: 47.5

#BSCB2
```{r}
chr2 = filter(df, Chrom==2)
head(chr2[order(abs(80.6e6-chr2$Position)),],n=10)
```
###Starting cM: 67
```{r}
head(chr2[order(abs(114.5e6-chr2$Position)),],n=10)
```
###Ending cM: 67.5

#BSCB4
```{r}
head(chr4[order(abs(82.7e6-chr4$Position)),],n=10)
```
###Starting cM: 55.6
```{r}
head(chr4[order(abs(140e6-chr4$Position)),],n=10)
```
###Ending cM: 57

##BSCB8
```{r}
chr8= filter(df, Chrom==8)
head(chr8[order(abs(125.1e6-chr8$Position)),],n=10)
```
###Starting cM: 61.5
```{r}
head(chr8[order(abs(145.6e6-chr8$Position)),],n=10)
```
###Ending cM: 67.7